# Practice

```{r}

ch5ds1 <- read.csv("../data/05/ch5ds1.csv")

# Range
range <- max(ch5ds1$Reaction.Time) - min(ch5ds1$Reaction.Time)

# Calculate mean
mean <- mean(ch5ds1$Reaction.Time, na.rm = T)

# Completing formula for variance 
sum <- 0
for (value in ch5ds1$Reaction.Time) {
  num <- (value - mean(ch5ds1$Reaction.Time))^2
  sum <- sum + num 
}
var <- sum / (length(ch5ds1$Reaction.Time) - 1)
# var
# var(ch5ds1$Reaction.Time) # checking result with built in function

# SD is the sqrt of Var
sd <-sqrt(var)

print(list(range, var, sd))
```


# Exercises 

5. For the following set of scores, compute the range, the unbiased and the biased standard deviations, and the variance. Do the exercise by hand.

```{r}

num_5 <- c(94, 86, 72, 69, 93, 79, 55, 88, 70, 93)

n <- length(num_5)

sum_mean <- 0

for (value in num_5) {
  sum_mean <- sum_mean + value
}

mean <- sum_mean / n

sum_var <- 0

for (value in num_5) {
  num <- (value - mean)^2
  sum_var <- sum_var + num
}

unbiased <- n - 1
biased <- n 

range <- max(num_5) - min(num_5)
var_unbiased <- sum_var / unbiased
sd_unbiased <- sqrt(var)

var_biased <- sum_var / biased
sd_biased <- sqrt(var_biased)

# Answers

print(paste0("Range: ", range))
print(paste0("Variance Unbiased: ", var_unbiased))
print(paste0("SD Unbiased: ", sd_unbiased))
print(paste0("Variance Biased: ", var_biased))
print(paste0("SD Biased: ", sd_biased))
```
6. In Question 5, why is the unbiased estimate greater than the biased estimate?

- Because the unbiased estimate takes a more conservative approach by reducing the sample size by 1. As behavioral scientists, we want to err on the side of caution. 

Using R, compute all of the measures of variability you can for height and weight.

12. How can you tell whether R produces a biased or an unbiased estimate of the standard deviation?

- By checking help for the function producing the estimate.

13. Compute the biased and unbiased values of the standard deviation and variance for the set of accuracy scores shown in Chapter 5 Data Set 4 (ch5ds4.csv). Use R if you can; otherwise, do it by hand. Which one is smaller and why?

```{r}

num_13 <- read.csv("../data/05/ch5ds4.csv")

mean <- mean(num_13$Accuracy)
n <- length(num_13$Accuracy)

sum_var <- 0

for (value in num_13$Accuracy) {
  num <- (value - mean)^2
  sum_var <- sum_var + num
}

unbiased <- n - 1
biased <- n 

var_unbiased <- sum_var / unbiased
sd_unbiased <- sqrt(var)

var_biased <- sum_var / biased
sd_biased <- sqrt(var_biased)

# Answers
print(paste0("Variance Unbiased: ", var_unbiased))
print(paste0("SD Unbiased: ", sd_unbiased))
print(paste0("Variance Biased: ", var_biased))
print(paste0("SD Biased: ", sd_biased))
```


14. On a spelling test, the standard deviation is equal to 0.94. What does this mean?

- This means that the average distance in scores from the mean for this sample is 0.94
